### 2023-09-19 manuscript/main.tex
Include N. Salamin comments

%! BibTeX Compiler = biber
%TC:ignore
\documentclass{article}
\usepackage{caption}
\usepackage{censor}
\usepackage{xcolor, colortbl}
\definecolor{BLUELINK}{HTML}{0645AD}
\definecolor{DARKBLUELINK}{HTML}{0B0080}
\definecolor{LIGHTGREY}{gray}{0.9}
\PassOptionsToPackage{hyphens}{url}
\usepackage[colorlinks=false]{hyperref}
% for linking between references, figures, TOC, etc in the pdf document
\hypersetup{colorlinks,
    linkcolor=DARKBLUELINK,
    anchorcolor=DARKBLUELINK,
    citecolor=DARKBLUELINK,
    filecolor=DARKBLUELINK,
    menucolor=DARKBLUELINK,
    urlcolor=BLUELINK
} % Color citation links in purple
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{naturalnames}{hyperref}

\usepackage{biorxiv}
\usepackage[backend=biber,eprint=false,isbn=false,url=false,giveninits=true, style=apa,date=year]{biblatex}
\DeclareDelimFormat{nameyeardelim}{\addcomma\space}
\addbibresource{references.bib}

\usepackage{url}
\usepackage{amssymb,amsfonts,amsmath,amsthm,mathtools}
\usepackage{lmodern}
\usepackage{xfrac, nicefrac}
\usepackage{bm}
\usepackage{listings, enumerate, enumitem}
\usepackage[export]{adjustbox}
\usepackage{graphicx}
\usepackage{bbold}
\usepackage{pdfpages}
\pdfinclusioncopyfonts=1
\usepackage{lineno}
\usepackage{tabu}
\usepackage{hhline}
\usepackage{multicol,multirow,array}
\usepackage{etoolbox}
\usepackage{booktabs}
\usepackage{makecell}
\usepackage{orcidlink}
\usepackage{marvosym}

% -- Defining colors:
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}% Definig a custom style:
\lstdefinestyle{mystyle}{
    backgroundcolor=\color{backcolour},
    basicstyle=\ttfamily\scriptsize\bfseries,
    breakatwhitespace=false,
    breaklines=true,
    captionpos=t,
    keepspaces=true,
    showspaces=false,
    showstringspaces=false,
    showtabs=false,
    tabsize=2
}% -- Setting up the custom style:
\lstset{style=mystyle}
\captionsetup[table]{hypcap=false}
\captionsetup[figure]{hypcap=false}

\newcommand{\defEqual}{\stackrel{\text{def}}{=}}
\newcommand{\Multiply}{\cdot}
\newcommand{\MultiplyMatrix}{\times}
\newcommand{\UniDimArray}[1]{\bm{#1}}
\newcommand{\BiDimArray}[1]{\bm{#1}}
\newcommand{\tr}{^{\intercal}}
\newcommand{\inv}{^{-1}}
\DeclareMathOperator{\E}{\mathbb{E}}
\DeclareMathOperator{\Var}{\text{Var}}
\DeclareMathOperator{\Cov}{\text{Cov}}
\newcommand{\QstFst}{Q$_\text{ST}$--F$_\text{ST}$}
\newcommand{\der}{\mathrm{d}}
\newcommand{\e}{\text{e}}
\newcommand{\Ne}{N_{\text{e}}}
\newcommand{\dn}{d_N}
\newcommand{\ds}{d_S}
\newcommand{\dnds}{\dn / \ds}
\newcommand{\pn}{\pi_N}
\newcommand{\ps}{\pi_S}
\newcommand{\pnps}{\pn / \ps}
\newcommand{\proba}{\mathbb{P}}
\newcommand{\pfix}{\proba_{\text{fix}}}
\newcommand{\Spi}{i}
\newcommand{\Spj}{j}
\newcommand{\NbrTaxa}{n}
\newcommand{\Time}{t}
\newcommand{\Trait}{P}
\newcommand{\Heritability}{h^2}
\newcommand{\HeritabilitySpi}{\Heritability_{\Spi}}
\newcommand{\MeanTrait}{\bar{\Trait_{\Time}}}
\newcommand{\VecTrait}{\UniDimArray{\bar{\Trait}}}
\newcommand{\Root}{0}
\newcommand{\RootTrait}{\Trait_{\Root}}
\newcommand{\VarPhy}{\Var \left[\MeanTrait\right]}
\newcommand{\VecZero}{\UniDimArray{0}}
\newcommand{\VecOne}{\UniDimArray{1}}
\newcommand{\Distance}{\BiDimArray{D}}
\newcommand{\DistanceMatrix}{\BiDimArray{\Distance}}
\newcommand{\MutationRate}{\mu}
\newcommand{\SubRate}{q}
\newcommand{\NbrLoci}{L}
\newcommand{\VarPhenotype}{V_{\Trait}}
\newcommand{\VarPhenotypeSpi}{V_{\Trait, \Spi}}
\newcommand{\VarGenetic}{V_{\mathrm{A}}}
\newcommand{\VarGeneticSpi}{V_{\mathrm{A}, \Spi}}
\newcommand{\VarEnv}{V_{\mathrm{E}}}
\newcommand{\VarMutation}{V_{\mathrm{M}}}
\newcommand{\GenArchi}{\NbrLoci \Multiply \E \left[ a^2 \right]}
\newcommand{\RateMut}{\sigma^2_{\mathrm{M}}}
\newcommand{\RateBetween}{\sigma^2_{\mathrm{B}}}
\newcommand{\RateWhithin}{\sigma^2_{\mathrm{W}}}
\newcommand{\RateWhithinSpi}{\sigma^2_{\mathrm{W}, \Spi}}
\newcommand{\VecRateWhithin}{\UniDimArray{\RateWhithin}}
\newcommand{\EstRateBetween}{\widehat{\RateBetween}}
\newcommand{\EstRateWhithin}{\widehat{\RateWhithin}}
\newcommand{\NI}{\rho}
\newcommand{\EstNI}{\widehat{\rho}}
\newcommand{\StdSelection}{\sigma}
\newcommand{\VarSelection}{\StdSelection^2}

% Tree
\newcommand{\Nbranch}{2 \NbrTaxa - 2}
\newcommand{\WishartPostDf}{2 \NbrTaxa + 1}
\newcommand{\Ntrait}{K}
\newcommand{\contrast}{\UniDimArray{C}}
\newcommand{\Covariancematrix}{\Sigma}
\newcommand{\CovarianceMatrix}{\BiDimArray{\Covariancematrix}}
\newcommand{\Precisionmatrix}{\Omega}
\newcommand{\PrecisionMatrix}{\BiDimArray{\Precisionmatrix}}
\newcommand{\Identitymatrix}{\BiDimArray{I}}
\newcommand{\brownian}{\mathcal{B}}
\newcommand{\Brownian}{\UniDimArray{\brownian}}
\newcommand{\Scattermatrix}{\BiDimArray{A}}
\newcommand{\Multivariate}{\UniDimArray{Z}}

\renewcommand{\baselinestretch}{1.5}
\renewcommand{\arraystretch}{1.2}
\linenumbers
\frenchspacing

% Including genomics, Inclusion of genomics, Linking genomics,
\title{Leveraging genomics to detect trait selection from within and between-species variations}
\rhead{\scshape Trait selection from within and between-species variations}

\author{
\large
\textbf{T. {Latrille}$^{1}$\orcidlink{0000-0002-9643-4668}, M. {Bastian}$^{2}$, T. {Gaboriau}$^{1}$\orcidlink{0000-0001-7530-2204}, N. {Salamin}$^{1}$\orcidlink{0000-0002-3963-4954}}\\
\normalsize
$^{1}$Department of Computational Biology, Université de Lausanne, Lausanne, Switzerland\\
$^{2}$Laboratoire de Biométrie et Biologie Evolutive, UMR5558, Université Lyon 1, Villeurbanne, France \\
\texttt{\href{mailto:thibault.latrille@ens-lyon.org}{thibault.latrille@ens-lyon.org}} \\
}
\author{~}
\begin{document}

\maketitle

% Abstract (≤ 250 words)
%TC:endignore
\begin{abstract}
    To determine whether a trait is neutrally evolving or under selection, methods have been developed using either within or between-species variation.
    However, methods using within-species variation do not integrate the changes at the macro-evolutionary scale.
    Conversely, current methods using between-species variation usually discard within-species variation, thus not accounting for processes at the micro-evolutionary scale.
    The main goal of this study is to define a neutrality index for a quantitative trait, by combining within- and between-species variation.
    This neutrality index integrates nucleotide polymorphism and divergence for normalizing trait variation.
    As such, it does not require estimation of population size nor of time of speciation for normalization.
    Our index can be used to seek deviation from the null model of neutral evolution, and test for diversifying selection.
    Applied to brain mass and body mass at the mammalian scale, we show that brain mass is under diversifying selection.
    Finally, we show that our test is not sensitive to the assumption that population sizes, mutation rates and generation time are constant across the phylogeny, and automatically adjust for it.
\end{abstract}

\keywords{Quantitative genetics \and Trait evolution \and Selection \and Phylogenetics \and Population genetics }

% Research Article (7500 words)
\section*{Introduction}\label{sec:introduction}

Determining whether a trait is neutrally evolving or under a particular regime of selection has been a long-standing goal in evolutionary biology.
Fundamentally, distinguishing neutral evolution from selection requires determining which selective regime is supported by the observed variation of traits or sequences.
The variation of phenotypes (traits) and genotypes (sequences) can be observed at different scales, across different development stages at the individual level, across different individuals and populations at the species level, and finally across different species at the phylogenetic level.
All these systems require different assumptions and methodologies, and the endeavor to determine the selective regime for a given trait has thus incorporated theories, methods, and developments across various fields of evolutionary biology such as quantitative genetics, population genetics, phylogenetics and comparative genomics~\parencite{lynch_genetics_1998, walsh_evolution_2018}.

Leveraging individual variations within the same species, Genome-Wide Association Studies (GWAS) in humans have shown that traits are mostly polygenic (many loci associated with a given trait), are also mostly pleiotropic (many traits associated with a given locus), additive and mainly under stabilizing selection~\parencite{simons_population_2018, sella_thinking_2019}.
Additionally, contrasting both trait and genetic differentiation across several populations, \QstFst methods have allowed to determine the selective regime and to quantify the strength of selection acting on a trait~\parencite{martin_multivariate_2008, leinonen_qst_2013}.
For example, in the context of the evolution of gene expression, studies have shown that stabilizing selection is largely dominant~\parencite{whitehead_neutral_2006, gilad_natural_2006.
However, selective regime acting on gene expression is still controversial, and neutral evolution is still debated~\parencite{signor_evolution_2018, price_detecting_2022}.

To disentangle neutral evolution and selection, trait evolution can also be observed at a larger time scale.
For example, change in mean trait value accumulates linearly with time of divergence from a sister species, and also proportionally to the trait variance~\parencite{lande_genetic_1980, turelli_heritable_1984}.
Gene expression exhibits a similar accumulation as divergence in expression accumulates faster for gene with large within-species variation~\parencite{khaitovich_neutral_2004}.
Altogether, both the trait variance and the evolution in mean value can be used to test for trait selection in a pair of species~\parencite{walsh_evolution_2018}.

% Phylogenetics (mean trait evolution)
Alternatively, by accounting for the underlying relationships between several species, the selective regime for a quantitative trait can also be tested at the phylogenetic scale~\parencite{felsenstein_phylogenies_1985}.
Under neutral evolution, the change in main trait value along a given branch of the tree is normally distributed, with a variance proportional to divergence time~\parencite{hansen_translating_1996}.
As a result, the mean trait value can be modeled as a Brownian process branching at every node of the tree~\parencite{hansen_translating_1996, harmon_phylogenetic_2018}.
Reconstructing the trait variation along the whole phylogeny as a Brownian process can thus constitute a null model of neutral trait evolution.
Some deviations from the Brownian process, typically Ornstein-Uhlenbeck processes, are thus often interpreted as a signature of stabilising selection~\parencite{catalan_drift_2019}.
Alternatively, a trend in the Brownian process (the tendency of a trait to evolve in a certain direction) is interpreted as a signature of directional selection at the phylogenetic scale~\parencite{silvestro_early_2019}.
However, studies have shown that such comparative approaches are subject to different biases~\parencite{harmon_phylogenetic_2018}.
First, a trait under stabilising selection for which the optimal trait value is also evolving as a Brownian process will not deviate from a Brownian process, and thus be wrongly classified as neutral~\parencite{hansen_translating_1996}.
In other words, the better fit of a Brownian process does not necessarily constitute proof of the neutral model.
Secondly, and contrarily, even under a neutral trait, the Ornstein-Uhlenbeck process might sometimes be statistically preferred over a Brownian process due to sampling artifacts~\parencite{silvestro_measurement_2015, cooper_cautionary_2016, price_detecting_2022}.
Those limitations, altogether with the use of mean trait estimates leaving out the variance in traits between individuals, easily generate misclassification of selection from methods at the phylogenetic scale.

% At the frontier Phylogenetics/Population-genetics
At the frontier between micro and macroevolution, comparative methods at the phylogenetic scale have acknowledged the importance of modeling within-species variation on top of changes in mean trait value to describe technical errors~\parencite{lynch_methods_1991, felsenstein_comparative_2008, hansen_interpreting_2012} or to scale the rate of change in mean trait value~\parencite{kostikova_bridging_2016, gaboriau_multiplatform_2020, gaboriau_exploring_2023}.
One notable exception used the ratio of between to within variation estimated for many traits, and compared to the average over all traits to seek deviation from this average as a signal of diversifying selection~\parencite{rohlfs_modeling_2014, rohlfs_phylogenetic_2015}.
The goal of this study is to determine the evolutionary regime of a quantitative trait through the articulation of the variance between and within species while setting the threshold for neutral, stabilising, and diversifying selection.
We propose a neutrality index for a quantitative trait articulating trait and nucleotide variation within and between species.
Importantly, our neutrality index also leverages nucleotide divergence and polymorphism in order to normalize trait variation at both scales, such that it does not require estimating population size (within-species) or speciation time (between-species).
For population geneticists, our study can be seen as the macro-evolutionary generalization of \QstFst methods to account for phylogenetic relationships between species.
For phylogeneticists, our study can be seen as an alternative to the EVE model~\parencite{rohlfs_modeling_2014, rohlfs_phylogenetic_2015}, where we set a clear threshold for neutral evolution by reviewing theoretical literature and leveraging species' nucleotide polymorphism and divergence.
% Basically, we are testing whether the variance of means is greater than the mean of variances

\section*{Materials and Methods}\label{sec:materials-and-methods}
\subsection*{Neutrality index for a quantitative trait}\label{subsec:neutrality-index-for-a-quantitative-trait}

Prior to developing our neutrality index, we review theoretical expectations for variations of quantitative traits and genomic sequences under neutral evolution for both within- and between-species variation.

\subsubsection*{Within-species trait variations}

For a given trait, the genetic architecture is mainly defined by the number of loci encoding the trait ($\NbrLoci$) and the random additive effect of a mutation on the trait ($a$).
New mutations are generating trait variance and the average effect of a mutation on the trait is $\RateMut = \NbrLoci \Multiply  \E [a^2]$.
At the individual level, the mutational variance ($\VarMutation$) is the rate at which new mutations contribute to the trait variance per generation.
As shown in \textcite{lande_quantitative_1979, lande_sexual_1980}, $\VarMutation$ is a function of the mutation rate per loci per generation ($\MutationRate$) and $\RateMut$:
\begin{gather}
    \VarMutation = 2 \MutationRate \Multiply \RateMut \label{eq:var-mutation}.
\end{gather}

While in an infinitesimal model mutations supply new genetic variants, random genetic drift depletes standing variation~\parencite{turelli_commentary_2017, barton_infinitesimal_2017, sella_thinking_2019}.
For a neutral trait at equilibrium between mutation and drift~\parencite{lynch_mutation_1998}, the additive genetic variance in a species ($\VarGenetic$) is a function of the mutational variance ($\VarMutation$) and the effective number of individuals in the population ($\Ne$):
\begin{align}
    \VarGenetic & =  2 \Ne \Multiply \VarMutation, \\
    & = 4 \Ne \Multiply \MutationRate \Multiply \RateMut \text{ from eq.~\ref{eq:var-mutation}}\label{eq:var-genetic}.
\end{align}

For any neutral genomic region of interest, the nucleotide diversity, $\pi$, is measured as the number of mutations segregating in the population divided by the length of the region.
Any segregating mutations will eventually reach fixation or extinction due to random genetic drift and $\pi$ is also at a balance between mutations and drift.
As shown in \textcite{tajima_statistical_1989}, $\pi$ is a function of the mutation rate per loci per generation ($\MutationRate$) and the effective population size ($\Ne$):
\begin{gather}
    \pi = 4 \Ne \Multiply \MutationRate \label{eq:genetic-diversity}.
\end{gather}

We define $\RateWhithin$ as the ratio of additive genetic variance of the trait ($\VarGenetic$) over $\pi$ of any neutral genomic region of interest.
This ratio allows to remove the effect of $\Ne$ and $\MutationRate$, which are parameters not related to the genetic architecture of the trait, giving $\RateWhithin$ as a proxy of $\RateMut$:
\begin{align}
    \RateWhithin & \defEqual \frac{\VarGenetic }{\pi}, \\
    & = \frac{4 \Ne \Multiply \MutationRate \Multiply \RateMut}{4 \Ne \Multiply \MutationRate} \text{ from eq.~\ref{eq:var-mutation} and~\ref{eq:genetic-diversity}}, \\
    & = \RateMut. \label{eq:rate-pop}
\end{align}

The additive genetic variance is also equal to the observed phenotypic variance ($\VarPhenotype$) multiplied by narrow-sense heritability ($\Heritability$; \parencite{hill_data_2008}), which leads to $\RateWhithin$ being a function of $\VarPhenotype$ and $\Heritability$:
\begin{align}
    \RateWhithin & = \frac{\Heritability \Multiply \VarPhenotype }{\pi }. \label{eq:rate-pheno-pop}
\end{align}

\subsubsection*{Between-species trait variations}

For a given species, we denote by $\MeanTrait$ the mean value of the trait across the individuals in the species at generation $\Time$.
If the trait is neutral and encoded by many loci as assumed by the infinitesimal model, $\MeanTrait$ evolves as a Brownian process~\parencite{felsenstein_phylogenies_1985, hansen_translating_1996}.
The variance of $\MeanTrait$ after $\Time$ generations, $\VarPhy$ is given by \parencite{hansen_translating_1996}:
\begin{align}
    \VarPhy & = \frac{\VarGenetic}{\Ne} \Multiply \Time \\
    & = 4 \Time \Multiply \MutationRate \Multiply \RateMut \label{eq:covar},\text{ from eq.~\ref{eq:var-genetic}},
\end{align}

Moreover, for any genomic region under neutral evolution, some mutations will eventually reach fixation due to random genetic drift, resulting in a substitution of a nucleotide at the species level.
The probability of fixation ($\pfix$) of a neutral mutation is $1/2\Ne$ \parencite{kimura_probability_1962}.
We can derive the substitution rate per generation $\SubRate$ as the number of mutations per generation ($2\Ne \Multiply \MutationRate$) multiplied by the probability of fixation for each newly arisen mutations $\pfix$~\parencite{mccandlish_modeling_2014}, giving:
\begin{align}
    \SubRate & = 2 \Ne \Multiply \MutationRate \Multiply \pfix, \\
    & = 2 \Ne  \Multiply \MutationRate  \Multiply \dfrac{1}{2\Ne}, \\
    & = \MutationRate. \label{eq:substitution-rate}
\end{align}
That is, if mutations are neutral, the rate of substitution within a genomic region equals the rate with which new mutations arise per generation for the same genomic region~\parencite{kimura_evolutionary_1968}.

After $\Time$ generations and assuming that no multiple substitutions occurred at the same site, the nucleotide divergence $d$, which is the fraction of the genomic region that generated a substitution, will be $\Time$ multiplied by the nucleotide substitution rate per generation ($\SubRate$):
\begin{align}
    d & = \Time \Multiply \SubRate \\
    & = \Time \Multiply \MutationRate \text{ from eq.~\ref{eq:substitution-rate}}. \label{eq:distance}
\end{align}
We define $\RateBetween$ as the variance in the mean trait value ($\VarPhy$) normalized by the nucleotide divergence of any neutral genomic region ($d$).
This ratio allows to remove the effect of $\Time$ and $\MutationRate$, which are parameters not related to the genetic architecture of the trade, giving $\RateBetween$ as another proxy of $\RateMut$:
\begin{align}
    \RateBetween & \defEqual \frac{\VarPhy}{4 d},\\
    & = \frac{4 \Time \Multiply \MutationRate \Multiply \RateMut}{4 \Time \Multiply \MutationRate}\text{ from eq.~\ref{eq:covar} and~\ref{eq:distance}}, \\
    & = \RateMut \label{eq:rate-phy}.
\end{align}

\begin{figure*}[!ht]
    \centering
    \includegraphics[width=\textwidth, page=1] {artworks/fig-input-output}
    \caption{
        Between species, the mean phenotypic trait value changes along the phylogeny, allowing estimating of between-species trait variation, $\EstRateBetween$, which is normalized by nucleotide divergence.
        Within-species, the genetic variance allows estimating of within-species trait variation, $\EstRateWhithin$, which is  normalized by nucleotide diversity.
        $\EstNI$ is the ratio of $\EstRateBetween$ over $\EstRateWhithin$.
        Under neutral evolution, $\EstNI$ is expected to be equal to one.
        Under diversifying selection, the trait is heterogeneous between species, but homogeneous within species, leading to $\EstNI$ greater than one.
        Under stabilising selection, the trait is homogeneous between species, leading to $\EstNI$ smaller than one.
        Importantly, the sequence from which nucleotide diversity and divergence are estimated should be neutrally evolving, but they are not necessarily linked to the quantitative trait under study, they allow for discarding the confounding effect on mutation rate diversity, population size and divergence time.
    }
    \label{fig:methods}
\end{figure*}

\subsubsection*{Neutrality index}

The variability between either individuals or species can be obtained for both quantitative traits and genomic sequences.
At the population level, the variability of the trait between individuals can be combined with the nucleotide diversity of any neutrally evolving genomic region to obtain $\RateWhithin$, which equals $\RateMut$ if the trait is neutrally evolving (see above).
At the phylogenetic level, the variability of the mean trait value between species can be combined with the nucleotide divergence of any neutrally evolving genomic region to obtain $\RateBetween$.
Similarly, $\RateBetween=\RateMut$ if the trait is neutrally evolving and the genetic architecture of the trait has not changed along the phylogeny.
We thus have, for a neutrally evolving trait:
\begin{gather}
    \RateWhithin = \RateBetween \text{ from eq.~\ref{eq:rate-pop} and~\ref{eq:rate-phy}}, \\
    \Rightarrow \NI \defEqual \frac{\RateBetween}{\RateWhithin} = 1.
\end{gather}
We define a neutrality index $\NI = \RateBetween / \RateWhithin$ that will equal $1$ for a trait evolving neutrally.
Both $\RateBetween$ and $\RateWhithin$ can be estimated using quantitative trait and genomic sequences within and between species, while neither the mutation rate ($\MutationRate$), nor the effective population size ($\Ne$) or time of divergence ($\Time$) need to be estimated.
Moreover, the sequence from which $\pi$ and $d$ are estimated should be neutrally evolving, but they are not necessarily linked to the quantitative trait under study.

\subsection*{Estimate}\label{subsec:estimate}

Based on the comparative framework accounting for phylogenetic inertia~\parencite{felsenstein_phylogenies_1985, omeara_testing_2006}, we provide a Maximum likelihood estimate for $\NI$ as well as a Bayesian estimate to derive posterior probabilities that the null model of neutrality (i.e.~$\NI = 1$) is rejected.

\subsubsection*{Maximum likelihood estimate}
At the phylogenetic scale, for $\NbrTaxa$ taxa in the tree, $\DistanceMatrix$ ($\NbrTaxa \times \NbrTaxa$) is the distance matrix computed from the branch lengths ($d$ as nucleotide divergence in units of substitution per site) and the topology of the phylogenetic tree.
The diagonal $\Distance_{\Spi,\Spi}$ represents the total distances from the root of the tree to each taxon ($\Spi$).
The off-diagonal elements ($\Distance_{\Spi,\Spj} = \Distance_{\Spj,\Spi}$) are the distances between the root and the most recent common ancestor of taxa $\Spi$ and $\Spj$.
The state $\RootTrait$ at the root of the tree for the trait can be estimated from the $\NbrTaxa \times 1$ vector of mean trait values $\VecTrait$ at the tips of the tree using maximum likelihood~\parencite{omeara_testing_2006}:
\begin{gather}
    \RootTrait = \left( \VecOne\tr \MultiplyMatrix \DistanceMatrix\inv \MultiplyMatrix \VecOne \right)\inv \Multiply \left( \VecOne\tr \MultiplyMatrix \DistanceMatrix\inv \MultiplyMatrix \VecTrait \right), \label{eq:estimated-root-trait}
\end{gather}
where $\VecOne$ is an $\NbrTaxa \times 1$ column vector of ones.

Finally, within-species variation $\EstRateBetween$ is estimated as~\parencite{omeara_testing_2006}:
\begin{gather}
    \EstRateBetween = \frac{1}{4}\frac{\left( \VecTrait -  \RootTrait \Multiply \VecOne \right)\tr \MultiplyMatrix \DistanceMatrix\inv \MultiplyMatrix \left( \VecTrait -  \RootTrait \Multiply \VecOne  \right)}{\NbrTaxa - 1}. \label{eq:estimated-rate-phy}
\end{gather}

For a given species $\Spi$ with inter-individual data available, additive genetic variance of a trait ($\VarGeneticSpi$) is the product of heritability ($\Heritability_{i}$) and phenotypic variance ($V_{\Trait, i}$).
The ratio of $\VarGeneticSpi$ over nucleotide diversity of neutrally evolving sequences ($\pi_{\Spi}$) is a sample estimate of $\RateWhithin$.
Averaged across all species, we obtain the estimate $\EstRateWhithin$ as:
\begin{gather}
    \EstRateWhithin = \dfrac{1}{\NbrTaxa}\sum_{i=1}^{\NbrTaxa}\frac{  \VarGeneticSpi}{ \pi_{i}} = \dfrac{1}{\NbrTaxa}\sum_{i=1}^{\NbrTaxa} \frac{  V_{\Trait, i} \Multiply \Heritability_{i}}{ \pi_{i}}. \label{eq:estimated-rate-pop}
\end{gather}

As depicted in fig.~\ref{fig:methods}, the neutrality index is estimated as:
\begin{gather}
    \EstNI = \frac{\EstRateBetween}{\EstRateWhithin}. \label{eq:estimated-NI}
\end{gather}

\subsubsection*{Bayesian estimate}

The Bayesian framework allows to obtain the posterior distribution the neutrality index ($\EstNI$) for a given trait.
Even though $\EstNI$ is estimated independently for each trait of interest in the maximum likelihood framework (previous section), here we generalize to $\Ntrait$ traits co-varying along the phylogeny using the \textit{BayesCode} software~\parencite{latrille_inferring_2021}.
Trait variation along the phylogeny is modeled as a $\Ntrait$-dimensional Brownian process $\Brownian$ ($1 \times \Ntrait$) starting at the root and branching along the tree topology~\parencite{huelsenbeck_detecting_2003, lartillot_phylogenetic_2011, lartillot_joint_2012, latrille_inferring_2021}.
The rate of change of the Brownian process is determined by the positive semi-definite and symmetric covariance matrix between traits $\CovarianceMatrix$ ($\Ntrait \times \Ntrait$).
The off-diagonal elements of $\CovarianceMatrix$ are the covariance between traits, and the diagonal elements are the variance of each trait, thus corresponding to $\EstRateBetween$ (see section~S\ref{subsec:multivariate-brownian-process}).
With an inverse Wishart distribution as the {prior} on the covariance matrix, the {posterior} on $\CovarianceMatrix$, conditional on $\brownian$ is also an invert Wishart distribution (see section~S\ref{subsec:sampling-the-covariance-matrix}).
$\Brownian$ is sampled using a Metropolis-Hastings algorithm (acceptance-rejection), and the posterior distribution of $\CovarianceMatrix$ is Gibbs sampled.
For each trait and each species, prior on heritability ($\Heritability$) for each species is set as a uniform distribution with user-defined boundaries.
Heritability and phenotypic variance for each trait is combined with nucleotide diversity to compute $\EstRateWhithin$ for each species, and is averaged out across species (as in eq.~\ref{eq:estimated-rate-pop}).
From $\EstRateWhithin$ and $\CovarianceMatrix$, the posterior distribution of $\EstNI$ (as in eq.~\ref{eq:estimated-NI}) is obtained for each trait.
The posterior distribution of $\EstNI$ thus allow to test for deviation from neutrality (see fig.~\ref{fig:methods}), for example by computing $\proba [\EstNI > 1 ]$ to test for evidence of diversifying selection and $\proba [\EstNI < 1 ]$ to test for evidence of stabilising selection.

\subsubsection*{Applicability to empirical data}

Our method assumes that the narrow-sense heritability ($\Heritability$) of a trait is known such as to estimate additive genetic variance ($\VarGenetic$) from phenotypic variance ($\VarPhenotype$) as $\VarGenetic = \Heritability \Multiply \VarPhenotype$.
Fortunately, if heritability is not known, the test for diversifying selection can still be performed, although it is underpowered.
Indeed, if the additive genetic variance is substituted by phenotypic variance, it is equivalent to assuming complete heritability ($\Heritability = 1$).
Because $\Heritability \leq 1$ by definition, we overestimate the within-species variation and thus underestimate $\EstNI$.
It is however possible to test for diversifying selection because testing for $\EstNI > 1$ while using phenotypic variance instead of additive genetic variance means that knowing the additive genetic variance would have only increased the evidence for diversifying selection.
Similarly, using the broad-sense heritability ($H^2$) instead of narrow-sense heritability ($\Heritability$) results in under-estimated $\EstNI$ since $\Heritability \leq H^2$.
Contrarily, the test for stabilising selection is invalid if $\EstNI$ is under-estimated.
Several assumptions made by our test might not hold on empirical data and their consequences on the neutrality index and the test that can be performed are shown in Table~\ref{table:assumptions}.

\subsection*{Simulation}\label{subsec:simulations}

% vG: 13.8, vP: 69.0, vE: 55.2, nbr_loci: 5000, a: 1.0, mut_rate: 1.38e-05, pop_size: 50, h2: 0.2
% pS = 0.0027577478571428568
% Tree length (dS) = 2.340579
% Tree length (My) = 1259.4490200000002
% Root age (My) = 98.9489413888889
% Root age (dS) = 0.18388820079995308
% nbr_sites_var = 27.577478571428568
% u = 1.3788739285714283e-05
% nbr_generations = 13336.114128321291
We tested the performance of our neutrality index ($\NI$) to detect selection on a quantitative trait using simulations.
We performed simulations under different selective regimes (neutral, stabilising, diversifying), different demographic histories (constant or fluctuating population size) and different evolution of mutation rate (constant or fluctuating).
Simulation are individual-based and follows a Wright-Fisher model with mutation, selection and drift for a diploid population including speciation along a predefined ultrametric phylogenetic tree (Fig.~\ref{fig:simulator}A\&B).
Each individual phenotypic value is the sum of genotypic value and an environmental effect.
The environmental effect was normally distributed with variance $\VarEnv$.
We assumed that the genotypic value was encoded by $\NbrLoci=5,000$ loci, with each locus contributing an additive effect that was normally distributed with standard deviation $a=1$ (Fig.~\ref{fig:simulator}A and section~S\ref{subsec:genotype-phenotype-map} for theoretical formulation).
We assumed a trait with a narrow-sense heritability of $\Heritability=0.2$ and computed the theoretical $\VarEnv$ accordingly (see section~S\ref{subsec:genotype-phenotype-map}).
Assuming a diploid panmictic population of size $\Ne=50$ at the root of the tree, and with non-overlapping generations, we simulated explicitly each generation along an ultrametric phylogenetic tree.
For each offspring, the number of mutations was drawn from a Poisson distribution with mean $2 \Multiply \MutationRate \Multiply \NbrLoci $, with mutation rate per generation $\MutationRate$.
From the empirical mammalian dataset (see next section), we computed an average nucleotide divergence from root to leaves of $0.18$ and average genetic diversity of $0.00276$.
We scaled parameters to fit plausible values for mammals.
We used a mutation rate of $\MutationRate=0.00276 / 4 \Ne = 1.38 \times 10^{-5}$ per generation per locus and a total of $\Time = 0.18 / 1.38 \times 10^{-5} = 13,500$ generations from root to leaves, and the number of generations along each branch was proportional to the branch length.

The changes in log-$\MutationRate$ and log-$\Ne$ along the lineages were both modeled by a geometric Brownian process ($\brownian \left(0, \sigma_{\MutationRate}=0.0086\right)$ and $\brownian \left(0, \sigma_{\Ne}=0.0086\right)$, meaning standard deviation of $0.0086 \Multiply \sqrt {13,500} = 1.0$ in log-space from root to leaves.
Additionally, the instant value of log-$\Ne$ was modeled as a independent sum of a geometric Brownian process and an Ornstein-Uhlenbeck process ($\text{OU} \left(0, \sigma_{\Ne}=0.1, \theta_{\Ne}=0.9\right)$).
The geometric Brownian motion accounted for long-term fluctuations (low rate of changes $\sigma_{\Ne}$ but unbounded), while the Ornstein-Uhlenbeck introduced short-term fluctuations (high rate of changes $\sigma_{\Ne}$ but bounded and mean-reverting).
The simulation started from an initial sequence at equilibrium at the root of the tree and, at each node, the process was split until it finally reached the leaves of the tree.
From a speciation process perspective, this is equivalent to an allopatric speciation over one generation.

Random genetic drift was introduced by re-sampling individuals at each generation, with each parent having a probability of being sampled that was proportional to its fitness ($W$).
Selection was modeled as a one-dimensional Fisher's geometric landscape, with the fitness of an individual being a monotonously decreasing function of the distance between the individual and optimal phenotype~\parencite{tenaillon_utility_2014,blanquart_epistasis_2016}.
More specifically, the fitness of an individual was given by $W = \e^{(\Trait - \lambda)^2/ \alpha}$, where $\Trait$ was the trait value of the individual, $\lambda=0.0$ was the optimal trait value, and $\alpha=0.02$ was the strength of selection.
Mutations were considered as a displacement of the phenotype in the multidimensional space.
Beneficial mutations moved the phenotype closer to the optimum, while deleterious mutations moved it further away.
Stabilizing selection was implemented by fixing the optimum phenotype to a single value ($\lambda=0.0$).
Diversifying selection was implemented by allowing the optimum phenotype to move along the phylogenetic tree as a geometric Brownian process~\parencite{hansen_stabilizing_1997} ($\lambda \sim \brownian \left(0, \sigma_{\lambda}=1.0\right)$).
Neutral evolution was implemented by fixing the fitness landscape ($W=1$), which meant that each individual had the same probability of being sampled at each generation.

Nucleotide diversity ($\pi$) was measured as the heterozygosity of neutral markers (not linked to the trait).
Nucleotide divergence ($d$) was measured as the number of substitutions per site of neutral markers (not linked to the trait) along branches of the phylogenetic tree.
The additive genetic variance was measured as phenotypic variance multiplied by heritability.
Heritability was estimated from the slopes of the regression of offspring's phenotypic trait values on parental phenotypic trait values~\parencite{lynch_genetics_1998} averaged over the last 10 simulated generations.
Heritability was thus not a given parameter of the simulations, but rather measured as it would be in empirical data.

\begin{figure*}[!ht]
    \centering
    \includegraphics[width=\textwidth, page=1] {artworks/fig-simulator}
    \caption{
        Wright-Fisher simulations with mutation, selection and drift.
        Left panel: For a given individual, the trait phenotypic value is the sum of genotypic value and a environmental effect (standard deviation $\VarEnv$).
        The trait's genotypic value is encoded by $\NbrLoci$ loci, with each locus contributing additively to the genotypic value.
        Parents are selected for reproduction according to their phenotypic value, with a probability proportional to their fitness.
        Mutations are drawn from a Poisson distribution, with each locus having a probability $\MutationRate$ to mutate.
        Drift is modeled by the resampling of parents.
        Right panel: example of a trait evolving along a phylogeny, with the mean phenotype (black line) and the variance of the trait genotypic value (gray area).
    }
    \label{fig:simulator}
\end{figure*}

\subsection*{Empirical dataset}\label{subsec:empirical-dataset}

We analyzed a dataset of body and brain masses in mammals.
The log-transformed values of body and brain masses were taken from \textcite{tsuboi_breakdown_2018}.
We removed individuals not marked as adults and split the data into males and females due to sexual dimorphism in body and brain masses.
We discarded species with only one representative sample.
The mammalian nucleotide diversity was obtained from the Zoonomia project~\parencite{genereux_comparative_2020}, with nucleotide divergence obtained on a set of neutral markers in \textcite{foley_genomic_2023}, and with nucleotide diversity measured as heterozygosity in \textcite{wilder_contribution_2023}.

We also analyzed a dataset of primate species, with the nucleotide variation obtained from \textcite{kuderna_global_2023} and the quantitative trait variation also from \textcite{tsuboi_breakdown_2018}, using the same filtering as for the mammalian dataset.
However, the primate nucleotide divergence was not obtained on a set of neutral markers as for the mammalian dataset, but across the whole genome.

\section*{Results}\label{sec:results}

\subsection*{Neutrality index}

For a neutral trait, the genetic architecture, meaning the number of loci encoding the trait and the average effect of a mutation on the trait, is formally related to both within and between-species variation of the trait.
We defined the neutrality index as $\NI = \RateBetween/\RateWhithin$, which equals $1$ for a neutral trait (see methods), suggesting that traits for which this relationship is not verified are putatively under selection.
Under stabilising selection, the variation between species is depleted because the mean trait value is maintained similar between different species, which leads to $\NI < 1$.
In contrast, under diversifying selection, the variation between species is inflated because species will have potentially different trait values~\parencite{hansen_stabilizing_1997}, which leads to $\NI > 1$.
Our neutrality index for a quantitative trait leverages the data for any number of species, and takes advantage of the signal over the whole phylogeny, while at the same time taking into account phylogenetic inertia and addressing the non-independence between-species (see fig.~\ref{fig:methods}).
This statistic can be obtained as a maximum likelihood estimate ($\EstNI$), from eq.~\ref{eq:estimated-rate-pop} and~\ref{eq:estimated-rate-phy}.
We devised a Bayesian estimate to obtain the posterior distribution of the neutrality index, and test for diversifying selection as $\proba [\EstNI > 1]$, and stabilising selection as $\proba [\EstNI < 1]$.

\begin{table*}[t!]
    \centering
    \begin{adjustbox}{width = 1.0\textwidth}
        \begin{tabular}{||l|l||c|c||c|c||}
            \hline
            Broken assumption                                       & Consequences                                       & $\EstRateWhithin$   & $\EstRateBetween$   & Test $\NI > 1$ & Test $\NI < 1$ \\ \hline \hline
            Trait encoded by few loci                        & Between-species trait variation is underestimated & --              & Underestimated & Conservative & Invalid  \\ \hline
            Sexual dimorphism                                & Within-species trait variation is overestimated   & Overestimated & -- & Conservative & Invalid  \\ \hline
            Inbreeding                                       & Nucleotide diversity ($\pi$) is underestimated    & Overestimated  & --              & Conservative & Invalid  \\ \hline
            Markers for polymorphism are negatively selected & Nucleotide diversity ($\pi$) is underestimated  & Overestimated & -- & Conservative & Invalid  \\ \hline
            Markers for polymorphism are positively selected & Nucleotide diversity ($\pi$) is underestimated  & Overestimated & -- & Conservative & Invalid  \\ \hline
            Markers for divergence are positively selected   & Nucleotide divergence ($d$) is overestimated & -- & Underestimated & Conservative & Invalid  \\ \hline
            Markers for polymorphism under balanced selection & Nucleotide diversity ($\pi$) is overestimated  & Underestimated & -- & Invalid & Conservative  \\ \hline
            Markers for divergence are negatively selected   & Nucleotide divergence ($d$) is underestimated & -- & Overestimated & Invalid & Conservative  \\ \hline
            Multiple nucleotide substitutions at the same locus & Nucleotide divergence ($d$) is underestimated & -- & Overestimated & Invalid & Conservative  \\\hline \hline
        \end{tabular}
    \end{adjustbox}
    \caption{Assumptions breaks and their consequences on the estimation of within-species variation ($\EstRateWhithin$), between-species variation ($\EstRateBetween$), and on the neutrality index $\NI = \EstRateBetween/\EstRateWhithin$.
    The two last columns indicate whether the test for diversifying selection ($\NI > 1$) and for stabilising selection $\NI < 1$ are conservative or invalid due to violated assumptions.
    }
    \label{table:assumptions}
\end{table*}

\subsection*{Results against simulations}\label{subsec:results-against-simulations}

The inference framework was first tested on independently simulated datasets matching an empirically relevant mammalian empirical regime (see Materials and Methods).
Under constant population size ($\Ne$) and constant mutation rate ($\MutationRate$) across the phylogenetic tree (fig.~\ref{fig:results-simulations}, top row), we found no false negative for simulations of stabilising ($\proba [\EstNI < 1] > 0.975$; blue in fig.~\ref{fig:results-simulations}) or diversifying ($\proba [\EstNI > 1] > 0.975$; red in fig.~\ref{fig:results-simulations}) selection.
For simulations under neutral evolution, 77\% of those were correctly identified ($0.025 \leq \proba [\EstNI > 1] \leq 0.975$; yellow in fig.~\ref{fig:results-simulations}), while 21\% and 2\% were wrongly detected as stabilising or diversifying selection, respectively.
Once we introduced fluctuating $\Ne$ and $\MutationRate$ (see fig.~\ref{fig:results-simulations}, bottom row), our ability to identify simulations under either diversifying or stabilising selection remained the same with all cases detected correctly.
For simulations under neutral evolution, 51\% of the simulations were correctly detected ($0.025 \leq \proba [\EstNI > 1] \leq 0.975$), while 49\% were detected as stabilising selection ($\proba [\EstNI < 1] > 0.975$).

\begin{figure*}[!ht]
    \centering
    \begin{minipage}{0.49\textwidth}
        \includegraphics[width=\textwidth, page=1] {artworks/cst_L5000.rho}
    \end{minipage}
    \begin{minipage}{0.49\textwidth}
        \includegraphics[width=\textwidth, page=1] {artworks/cst_L5000.pvalues}
    \end{minipage}
    \begin{minipage}{0.49\textwidth}
        \includegraphics[width=\textwidth, page=1] {artworks/fluNe_L5000.rho}
    \end{minipage}
    \begin{minipage}{0.49\textwidth}
        \includegraphics[width=\textwidth, page=1] {artworks/fluNe_L5000.pvalues}
    \end{minipage}
    \caption{
        $10,000$ replicate simulations of trait evolution along a phylogeny under different selection regimes.
        Traits simulated under stabilising selection (blue), under a neutral evolution (yellow), and under a moving optimum (red).
        Histogram of estimated the ratio of between-species trait variation ($\EstRateBetween$) over within-species trait variation $\EstRateWhithin$, as $\EstNI = \EstRateBetween / \EstRateWhithin$ estimated from each simulated data (left) and the histogram of probabilities that the estimated $\EstNI$ is greater than $1$ (right).
        Effective population size ($\Ne$) and mutation rate ($\MutationRate$) are either constant (top row), or fluctuating as a Brownian process along the phylogeny (bottom row).
    }
    \label{fig:results-simulations}
\end{figure*}

\begin{table*}[t!]
    \centering
    \begin{adjustbox}{width = 0.75\textwidth}
        \begin{tabular}{||l|l|l|c|c|c|c|c||}
        \toprule
        Dataset & Trait & $\Heritability$ & Sex & $\NbrTaxa$ & $\EstNI$ & $95\%$ CI for $\EstNI$ & $\proba [\EstNI > 1 ]$ \\ \hline
        \midrule
        Mammals & Body mass & 1.0 & \Male & 36 & 0.340 & 0.217-0.523 & 0.000 \\ \hline
        Mammals & Body mass & 1.0 & \Female & 26 & 0.277 & 0.160-0.490 & 0.000 \\ \hline
        Mammals & Body mass & $\mathcal{U}(0.2, 0.4)$ & \Male & 36 & 1.124 & 0.721-1.754 & 0.635 \\ \hline
        Mammals & Body mass & $\mathcal{U}(0.2, 0.4)$ & \Female & 26 & 0.936 & 0.523-1.715 & 0.324 \\ \hline \hline
        Mammals & Brain mass & 1.0 & \Male & 36 & 1.351 & 0.851-2.173 & 0.877 \\ \hline
        Mammals & Brain mass & 1.0 & \Female & 26 & 1.727 & 0.991-2.938 & 0.972 \\ \hline
        Mammals & Brain mass & $\mathcal{U}(0.2, 0.4)$ & \Male & 36 & 4.527 & 2.831-7.091 & 1.000 \\ \hline
        Mammals & Brain mass & $\mathcal{U}(0.2, 0.4)$ & \Female & 26 & 6.001 & 3.288-10.941 & 1.000 \\ \hline \hline
        Primates & Body mass & 1.0 & \Male & 71 & 0.558 & 0.401-0.784 & 0.000 \\ \hline
        Primates & Body mass & 1.0 & \Female & 65 & 0.389 & 0.278-0.547 & 0.000 \\ \hline
        Primates & Body mass & $\mathcal{U}(0.2, 0.4)$ & \Male & 71 & 1.875 & 1.288-2.695 & 1.000 \\ \hline
        Primates & Body mass & $\mathcal{U}(0.2, 0.4)$ & \Female & 65 & 1.296 & 0.899-1.821 & 0.914 \\ \hline \hline
        Primates & Brain mass & 1.0 & \Male & 71 & 1.929 & 1.395-2.616 & 1.000 \\ \hline
        Primates & Brain mass & 1.0 & \Female & 65 & 1.950 & 1.399-2.790 & 1.000 \\ \hline
        Primates & Brain mass & $\mathcal{U}(0.2, 0.4)$ & \Male & 71 & 6.479 & 4.658-8.944 & 1.000 \\ \hline
        Primates & Brain mass & $\mathcal{U}(0.2, 0.4)$ & \Female & 65 & 6.522 & 4.664-9.294 & 1.000 \\
        \bottomrule
        \end{tabular}
    \end{adjustbox}
    \caption{
    Test of diversifying selection on a mammal and a primate dataset, by splitting males (\Male) and females (\Female).
    Traits considered are body mass or brain mass (log-transformed).
    Heritability ($\Heritability$) is either assumed complete ($\Heritability=1.0$) or uniformly distributed between 20\% and 40\%  ($\Heritability \sim \mathcal{U}(0.2, 0.4)$).
    $\NbrTaxa$ is the number of species in the dataset.
    $\EstNI$ is the posterior estimate of our neutrality index, with the $95\%$ credible interval (CI) for $\EstNI$ also computed.
    $\proba [\EstNI > 1 ]$ is the estimated posterior probability of diversifying selection.
    }
    \label{table:empirical}
\end{table*}

\subsection*{Results on empirical data}\label{subsec:results-on-empirical-data}

For mammalian body and brain mass, we obtained male (\Male) and female (\Female) trait variation.
Combined with nucleotide diversity and divergence, we estimated $\EstNI$ and posterior probabilities of diversifying selection under different assumptions for trait heritability as shown in table~\ref{table:empirical}.
Assuming complete heritability, brain mass was found to be under diversifying selection with posterior probabilities of $0.0$ for both males and females.
If we assumed that heritability ($\Heritability$) of body mass was uniformly distributed between 20\% and 40\%~\parencite{hu_bringing_2022}, posterior probabilities of diversifying selection became $0.635$ for males and $0.324$ for females.
Mammalian brain mass was found to be under diversifying selection with posterior probabilities of $0.877$ for males and $0.972$ for females when complete heritability was assumed.
Assuming a uniform distribution between 20\% and 40\% for heritability led to posterior probabilities of diversifying selection of $1.0$ for both males and females.

We also analyzed a similar dataset for body mass focussing this time only at Primates (see table~\ref{table:empirical}).
For primate body mass, we found posterior probabilities of diversifying selection of $1.0$ for males and $0.914$ for females when assuming a uniform distribution for the heritability of body mass between 20\% and 40\%,
Assuming complete heritability of body mass did not change the posterior probability for males, but increased the one for female to $1.0$.
Evidence for diversifying selection on body mass was therefore more pronounced in Primates than in Mammals.
However, the genetic markers used to normalize trait variance with divergence were not necessarily neutral, which could create spurious false positives by artificially inflating $\EstNI$ (see table~\ref{table:assumptions} and methods).

\section*{Discussion}\label{sec:discussion}

% Summary of the method and results
In this study, we proposed a neutrality index for a quantitative trait that can be used within a statistical framework to test for selection.
Our neutrality index for a trait, $\NI$, is the ratio of the normalized between- to within-species variation and it allows the identification of the evolutionary regime of a quantitative trait.
At the phylogenetic scale, trait variation between species is normalized by sequence divergence obtained from a neutral set of markers.
Similarly, trait variation within species is normalized by sequence polymorphism obtained from a neutral set of markers.
Estimated $\EstNI$ can be compared against the null hypothesis of neutrality by testing for deviation from the value of $1.0$.
Technically, the neutrality index is estimated either as a maximum likelihood point estimate, or as a mean posterior estimate from a Bayesian implementation (see section~S\ref{sec:implementation}) also giving access to the posterior credible interval allowing to test for a departure from a neutrally evolving trait (e.g. $ \proba [ \EstNI > 1 ]$).
We tested our statistical procedure against simulated data and showed that for simulations under diversifying selection (test of $\EstNI > 1$) or under stabilising selection (test of $\EstNI < 1$) our test is able to correctly detect those regimes of selection.
However, for simulations of a neutral trait, our test detects frequently a spurious signal of stabilising selection ($\EstNI < 1$).
Thus, we argue that our test should be used to detect diversifying selection, but has low accuracy to detect stabilising selection due to many false positives.

Based on these results, we argue that our method is a significant improvement over current methods to detect selection acting on a trait at the phylogenetic scale.
Current methods relying on evolution of the mean trait value between species also tend to statistically prefer a model of stabilising selection over a Brownian process when the trait is neutral~\parencite{silvestro_measurement_2015, cooper_cautionary_2016, price_detecting_2022}.
Thus our method does not provide improvement to detect stabilising selection at the phylogenetic scale.
However, to detect diversifying selection, current methods modelling mean trait value are limited by the theory behind~\parencite{harmon_phylogenetic_2018}
Namely that under diversifying selection, mean trait value will not deviate from a Brownian process, and thus can not be distinguished from neutral evolution~\parencite{hansen_translating_1996}.
For example, test for neutrality concluded that the expression level of the majority of genes are under a neutral regime~\parencite{catalan_drift_2019}.
Our diversity index allows to discriminate the alternative model of diversifying selection from the neutral case, inviting us to revisit these studies.
Leveraging the ratio of between-species over within-species variations, methods have compared this ratio across a pool of traits~\parencite{rohlfs_modeling_2014, rohlfs_phylogenetic_2015}.
This strategy allows to identify outlier traits putatively under diversifying selection, but does not allow to test for selection on a single trait and requires independent confirmation~\parencite{rohlfs_phylogenetic_2015, gillard_comparative_2021}.
Instead, our procedure can be applied to a single trait, estimating the neutrality index and giving a statistical test for departure from the null model of neutral evolution for a single test.
Again, our diversity index can thus allow us to revisit these studies, assuming we have access to genomic datasets to estimate nucleotide divergence and polymorphism.

We argue that the main novelty of this study is being able to use nucleotide divergence and polymorphism to normalize trait variation between and within species.
In the context of within species variation but across several populations, \QstFst test developed from trait and sequence allow to test for selection of a trait~\parencite{martin_multivariate_2008, leinonen_qst_2013}.
Similarly, in our test, the genetic sequences from which nucleotide divergence and polymorphism are estimated should be neutrally evolving, but they are not necessarily linked to the quantitative trait under study.
Nucleotide variation allow to normalize for diversity driven by confounding factors such as population sizes ($\Ne$), mutation rates ($\MutationRate$) and generation time~\parencite{hansen_translating_1996, harmon_phylogenetic_2018}.
Thus our test avoids to estimate them and also avoids to estimate divergence time which is otherwise necessary~\parencite{walsh_evolution_2018}.
But importantly, by normalising with sequence variation, we also show using simulated data that our test is not sensitive to the assumption that $\Ne$, $\MutationRate$ and generation time are constant across the phylogeny, an unmet assumption empirically~\parencite{bergeron_evolution_2023, wilder_contribution_2023}.
Indeed, under the neutral case of evolution, changes in $\Ne$, $\MutationRate$ and generation time impacts similarly trait and sequence variation.
Thus the normalisation by nucleotide divergence and polymorphism automatically absorbs long-term and short-term changes in $\Ne$, $\MutationRate$ and generation time which cancels out in the ratio of trait variation $\EstNI$.

% How is our test related to other tests of selection
Even though our test is developed for a quantitative trait, analogies with other tests of selection developed for molecular sequences also provide insight into its behavior.
First, we acknowledge that our test took inspiration from the \textcite{mcdonald_adaptative_1991} test devised for protein-coding DNA sequences, where synonymous mutations are used to determine the neutral expectation, and the inflation of divergence is compared to polymorphism within species.
Second, because $\NI$ is compared to 1, our test ultimately bears analogy to the codon-based test of selection, where the ratio of non-synonymous to synonymous substitutions ($\dnds$) is compared to 1~\parencite{goldman_codonbased_1994, muse_likelihood_1994}.
As $\dnds < 1$ is interpreted as purifying selection acting on the protein, $\NI < 1$ is interpreted as stabilising selection acting on the trait.
Similarly, the interpretation of adaptation for $\dnds > 1$ is analogous to diversifying selection for $\NI > 1$.
With this analogy in mind, we can leverage the vast bibliography discussing and interpreting results of these tests and their pitfalls~\parencite{nielsen_molecular_2005, anisimova_investigating_2009, jensen_importance_2019}.
First, not rejecting the neutral null model of $\NI = 1$ does not necessarily imply that the trait is effectively neutral, since diversifying and stabilising selection could compensate each other resulting in $\NI = 1$, analogously to $\dnds=1$ under a mix of adaptation and purifying selection~\parencite{nielsen_molecular_2005}.
Second, empirical evidence for $\NI < 1$ does not rule out diversifying selection, but rather that this diversifying selection is not strong enough to overcome the stabilising selection, similarly to strong purifying selection resulting $\dnds < 1$ even though those genes and sites are under adaptation~\parencite{latrille_genes_2023}.
By explicitly modeling stabilising selection as a moving optimum, it would theoretically allow to tease out the effect of diversifying and stabilising selection in the context of quantitative traits to obtain a statistically more powerful test.

% Pitfalls of the method
In the context of detecting diversifying selection, we argue that the main drawback of our method is that the additive genetic variance of the trait is required instead of the phenotypic variance.
If phenotypic variance is used instead of additive genetic variance to estimate $\EstNI$, meaning that we assumed complete heritability, the neutrality index $\EstNI$ is ultimately under-estimated.
Similarly, using broad-sense heritability instead of narrow-sense heritability results in under-estimated $\EstNI$.
In such context, the test of stabilising selection ($\EstNI < 1]$) is statistically invalid.
However, the test of diversifying selection ($\EstNI > 1$) is underpowered although not invalided, meaning that absence of evidence would not be evidence of absence.
As an example, even though we assumed complete heritability for brain mass, we uncovered diversifying selection in mammals since $\EstNI > 1$.
However, at the primate scale, the evidence for $\EstNI > 1$ does not necessarily imply that the brain mass is evolving under diversifying selection since the markers used for nucleotide divergence are not neutral, which can lead to a spurious $\EstNI > 1$ (see table~\ref{table:assumptions}).

% What can and should be improved
The development of our neutrality index is also based on several assumptions that could be relaxed in future studies.
First, we cannot predict the behaviour of our test in the context of population structures, gene flow and introgression.
These factors should be thoroughly investigated using simulations.
Second, loci are assumed to contribute additively to the phenotype.
Although the effects of dominance and epistasis is typically weak compared the additive effects on the phenotype, their influence shall be assessed~\parencite{hill_data_2008, crow_epistasis_2010}.
Third, the genetic architecture of the trait is assumed to be constant across the phylogeny, it might actually be variable among individuals and species~\parencite{tung_genetic_2015, huber_conservatism_2015}.
Such assumption can theoretically be relaxed and changes in genetic architecture along the phylogeny could jointly be estimated~\parencite{arnold_understanding_2008, hohenlohe_mipod_2008, kostikova_bridging_2016, gaboriau_multiplatform_2020}.
Also, our Bayesian estimation could integrate uncertainty coming from the estimation of genetic variation, using sequences as input instead of estimated values of nucleotide diversity and divergence.

% What can our test bring to the field
From an empirical point of view, even though our method requires integrating genomic and trait variation, we also showed that our method can be applied to the empirical dataset with the illustrative example of mammals' brain and body mass.
Because our test is also based on several assumptions that might not hold on empirical data, we also provide a table containing the main assumptions and their consequences on the neutrality index and the test that can be performed (see table~\ref{table:assumptions}).
Altogether, our study provides a statistical framework to test for diversifying selection acting on a quantitative trait while integrating the trove of genomic data available both within and between species, and we argue that it is a promising tool to investigate the evolution of quantitative traits.

%TC:ignore
\printbibliography

\newpage

\part*{Supplementary materials}
\renewcommand{\thetable}{S\arabic{table}}
\renewcommand{\thefigure}{S\arabic{figure}}
\setcounter{figure}{0}
\setcounter{table}{0}
\setcounter{section}{0}

\renewcommand{\baselinestretch}{1.0}\normalsize
\tableofcontents
\renewcommand{\baselinestretch}{1.5}\normalsize

\newpage
\section{Genetic architecture of the trait}\label{sec:simulator}

\subsection{Genotype-phenotype map}\label{subsec:genotype-phenotype-map}

\begin{itemize}
    \item $\NbrLoci$ is the number of loci encoding the trait.
    \item $a_l \sim \mathcal{N}(0,a^2)$ is the effect of a mutation on the trait at locus $l \in \{1, \hdots, \NbrLoci\}$.
    \item $\Ne$ is the effective number of individuals.
    \item $g_{i,l} \in \{0, 1, 2\}$ is the genotypic value at locus $l$ for individual $i \in \{1, \hdots, \Ne\}$.
    \item $G_i = \sum_{l=1}^{\NbrLoci} a_l \times g_{i,l}$ is the genotypic value for individual $i$.
    \item $\xi_i \sim \mathcal{N}(0, \VarEnv)$ is the effect of environment on the trait for individual i.
    \item $\Trait_i = G_i + \xi_i$ is the phenotype for individual $i$.
\end{itemize}

\begin{center}
    \captionof{figure}{summary of trait's genetic architecture.}
    \includegraphics[width=0.6\textwidth, page=1] {artworks/fig-simulator-model}
    \label{fig:simulator-summary}
\end{center}

within-species, the mean ($\bar{G}$) and variance ($\VarGenetic$) of the genotype are:
\begin{equation}
    \bar{G} = \frac{1}{\Ne}\sum_{i=1}^{\Ne} G_i  \text{\quad and \quad} \VarGenetic = \frac{1}{\Ne}\sum_{i=1}^{\Ne}\left(G_i - \bar{G} \right)^2\label{eq:simu-genotype}
\end{equation}
The theoretical additive genetic variance ($\VarGenetic$) is a function of the number of loci ($\NbrLoci$) and the effect of a mutation ($a$) as:
\begin{equation}
    \VarGenetic = 4 \Ne \Multiply \MutationRate \Multiply \NbrLoci \Multiply a^2 \label{eq:simu-var-genetic}
\end{equation}

The mean ($\bar{\Trait}$) and variance ($\VarPhenotype$) of the phenotype are:
\begin{equation}
    \bar{\Trait} = \frac{1}{\Ne}\sum_{i=1}^{\Ne} \Trait_i \text{\quad and \quad} \VarPhenotype = \frac{1}{\Ne}\sum_{i=1}^{\Ne}\left(\Trait_i - \bar{\Trait} \right)^2 \label{eq:simu-between}
\end{equation}

Heritability ($\Heritability$) is defined as:
\begin{equation}
    \Heritability = \frac{\VarGenetic}{\VarPhenotype} = \frac{\VarGenetic}{\VarGenetic + \VarEnv}\label{eq:simu-heritability}
\end{equation}
Altogether, effective population size ($\Ne$), the number of loci ($\NbrLoci$) and the effect of a mutation ($a$), we can compute the variance of the environment ($\VarEnv$) that is required to reach a given heritability ($\Heritability$) as:
\begin{equation}
    \VarEnv = \VarGenetic \Multiply \left( \frac{1}{\Heritability} - 1 \right) = 4 \Ne \Multiply \MutationRate \Multiply \NbrLoci \Multiply a^2 \Multiply \left( \frac{1}{\Heritability} - 1 \right) \label{eq:simu-var-env}
\end{equation}

\newpage
\section{Bayesian estimate}\label{sec:bayesian-estimate}

\subsection{Multivariate Brownian process}\label{subsec:multivariate-brownian-process}
Here we generalize to $\Ntrait$ traits evolving along the phylogeny and are correlated between them.
Their variation along the phylogeny is modeled as a $\Ntrait$-dimensional Brownian process $\Brownian$ ($1 \times \Ntrait$) starting a the root and branching along the tree topology.
The rate of change of the Brownian process is determined by the positive semi-definite and symmetric covariance matrix between traits $\CovarianceMatrix$ ($\Ntrait \times \Ntrait$).
Along branch $j$ with length $d_{j}$, the Brownian process start at the ancestral node $\mathcal{A}(j)$ with value $\Brownian(\mathcal{A}(j))$, and ends at node $\mathcal{R}(j)$  with value $\Brownian(\mathcal{R}(j))$.
The independent contrast $\contrast_{j}$ defined as change in trait along the branch normalized by $\sqrt {d_{j}}$ is a multivariate Gaussian:
\begin{equation}
    \label{eq:DistribBrownian}
    \contrast_{j} = \frac{\Brownian_{\mathcal{R}(j)} - \Brownian_{\mathcal{A}(j)} }{\sqrt {d_{j}}} \sim \mathcal{N}\left(\VecZero, \CovarianceMatrix \right).
\end{equation}

\subsection{Sampling the covariance matrix}\label{subsec:sampling-the-covariance-matrix}
From the independent contrast at each branch of the tree ($\contrast_{j}$), we can define the $\Ntrait \times \Ntrait$ scatter matrix, $\Scattermatrix$, as:
\begin{equation}
    \Scattermatrix = \sum\limits_{j=1}^{\Nbranch} \contrast_{j} \MultiplyMatrix \left[\contrast_{j}\right]\tr\label{eq:bayes-scatter},
\end{equation}
where $\Nbranch$ is the number of branches in the tree and $\NbrTaxa$ the number of taxa.

The {prior} on the covariance matrix is an inverse Wishart distribution, with $\Ntrait + 1$ degrees of freedom:
\begin{equation}
    \label{eq:Distribcovariance}
    \CovarianceMatrix \sim \text{Wishart}^{-1} (\Identitymatrix, \Ntrait + 1).
\end{equation}
By Bayes theorem, the {posterior} on $\CovarianceMatrix$, conditional on a particular realization of $\Brownian$ (and thus of $\contrast$) is an invert Wishart distribution, of parameter $\Identitymatrix + \Scattermatrix$ and with $\WishartPostDf$ degrees of freedom.
\begin{equation}
    \CovarianceMatrix \sim \text{Wishart}^{-1}\left( \Identitymatrix + \Scattermatrix, \WishartPostDf\right)\label{eq:bayes-posterior}
\end{equation}
This invert Wishart distribution can be obtained by sampling $\WishartPostDf$ independent and identically distributed multivariate normal random variables $\Multivariate_{k}$ defined by
\begin{equation}
    \Multivariate_{k} \sim \mathcal{N} \left( \VecZero, \left[ \Identitymatrix + \Scattermatrix\right]^{-1} \right).\label{eq:bayes-multivariate}
\end{equation}
And from these multivariate samples, $\CovarianceMatrix$ is Gibbs sampled as:
\begin{equation}
    \CovarianceMatrix = \left( \sum\limits_{k=1}^{\WishartPostDf} \Multivariate_{k} \MultiplyMatrix  \left[\Multivariate_{k} \right] \tr \right)\inv \label{eq:bayes-gibbs}
\end{equation}

\newpage
\section{Bayesian and Maximum-likelihood implementation}\label{sec:implementation}

Implementation is included within the \textit{BayesCode} software, available at \censor{\url{https://github.com/ThibaultLatrille/bayescode}}.

\subsection{Data formatting}\label{subsec:data-formatting}

Running the analysis on your dataset and compute posterior probabilities requires three files:
\begin{enumerate}
    \item A phylogenetic tree in newick format, with branch lengths in number of substitutions per site (neutral markers)
    \item A file containing the mean trait values for each species.
    \item A file containing the variation within-species for each trait and the genetic variation within-species (neutral markers).
\end{enumerate}

\subsubsection{Phylogenetic tree}\label{subsubsec:phylogenetic-tree}

The phylogenetic tree must be in newick format, with branch lengths in number of substitutions per site (neutral markers).

\subsubsection{Mean trait for each species}\label{subsubsec:mean-trait-for-each-species}

The file containing mean trait values for each species must be in a tab-delimited file with the following format:
\begin{center}
    \begin{adjustbox}{width = 0.35\textwidth}
        \begin{tabular}{|l|c|c|}
            \hline
            TaxonName            & Body\_mass & Brain\_mass \\
            \hline
            Panthera\_tigris     & 12.26      & 5.676       \\
            Pithecia\_pithecia   & 7.256      & 3.436       \\
            Colobus\_angolensis  & 9.176      & 4.284       \\
            Saimiri\_boliviensis & 6.845      & 3.279       \\
            $\vdots$             & $\vdots$   & $\vdots$    \\
            \hline
        \end{tabular}\label{tab:trait-mean}
    \end{adjustbox}
\end{center}

The columns are:
\begin{itemize}
    \item \emph{TaxonName}: the name of the taxon matching the name in the alignment and the tree.
    \item As many columns as traits, without spaces or special characters in the trait.
    \item The values can be \texttt{NaN} to indicate that the trait is not available for that taxon.
\end{itemize}

\newpage
\subsubsection{Trait variation for each species}\label{subsubsec:trait-variation-for-each-species}

The file containing trait variation for each species must be in a tab-delimited file with the following format:
\begin{center}
    \begin{adjustbox}{width = 1.0\textwidth}
        \begin{tabular}{|l|c|c|c|c|c|c|}
            \hline
            TaxonName            & Nucleotide\_diversity & Body\_mass\_variance & Body\_mass\_heritability & Brain\_mass\_variance & Brain\_mass\_heritability \\
            \hline
            Pithecia\_pithecia   & 0.0016                & 0.22871              & 0.2                      & 0.00737               & 0.2                       \\
            Colobus\_angolensis  & 0.0017                & 0.00393              & 0.2                      & 0.00416               & 0.2                       \\
            Saimiri\_boliviensis & 0.0013                & 0.00022              & 0.2                      & 0.00045               & 0.2                       \\
            Pygathrix\_nemaeus   & 0.0016                & 0.00347              & 0.2                      & 0.00097               & 0.2                       \\
            $\vdots$             & $\vdots$              & $\vdots$             & $\vdots$                 & $\vdots$              & $\vdots$                  \\
            \hline
        \end{tabular}
        \label{tab:trait-variance}
    \end{adjustbox}
\end{center}

\begin{itemize}
    \item \emph{TaxonName}: the name of the taxon matching the name in the alignment and the tree.
    \item \emph{Nucleotide\_diversity}: the nucleotide diversity within-species (neutral markers), cannot be \texttt{NaN}.
    \item As many columns as traits, without spaces or special characters in the trait.
    \item \emph{TraitName\_variance}: the phenotypic variance of the trait within-species, can be \texttt{NaN} to indicate that the trait variance is not available for that taxon.
    \item \emph{TraitName\_heritability} (optional): the heritability of the trait within-species, between 0 and 1, cannot be \texttt{NaN}.
    \item The columns with the suffix \texttt{\_variance} and \texttt{\_heritability} are repeated for each trait.
    \item \emph{TraitName\_heritability\_lower} (optional): the lower bound of the heritability of the trait within-species, between 0 and 1, cannot be \texttt{NaN}.
    \item \emph{TraitName\_heritability\_upper} (optional): the upper bound of the heritability of the trait within-species, between 0 and 1, cannot be \texttt{NaN}.
    \item If the columns with the suffix \texttt{\_heritability\_lower} and \texttt{\_heritability\_upper} are present, the heritability is randomly drawn from a uniform distribution between the lower and upper bounds.
    \item If the columns with the suffix \texttt{\_heritability} is present, it is taken as is.
    \item If the additive genetic variance (instead of phenotypic variance) is available for a trait, the heritability can be omitted and will automatically be set to 1.0.
\end{itemize}

\newpage
\subsection{Bayesian estimation}\label{subsec:running-nodetraitsand-readnodetraits}

The executable \texttt{nodetraits} from \textit{BayesCode} is used to run the Bayesian estimation of the model, and the executable \texttt{readnodetraits} is used to read the results.

Assuming that the file \texttt{data/body\_size/mammals.male.tsv} contains the mean trait values for each species, the file \texttt{data/body\_size/mammals.male.var\_trait.tsv} contains the variation within-species for each trait and the genetic variation within-species (neutral markers), and the file \texttt{data/body\_size/mammals.male.tree} contains the phylogenetic tree, the following commands are used to run the model and read the results.

\subsubsection{Running the model}\label{subsubsec:running-nodetraits}
\texttt{nodetraits} is run with the following command:
\begin{lstlisting}[language = sh,label={lst:nodetraits-run}]
nodetraits  --until 2000
            --tree data/body_size/mammals.male.tree
            --traitsfile data/body_size/mammals.male.tsv
            run_mammals_male
\end{lstlisting}

\subsubsection{Reading the results}\label{subsubsec:reading-the-results}
Once the model has ran, the chain \texttt{run\_mammals\_male} is used to compute the posterior distribution of the ratio of between-species variation over within-species variation with \texttt{readnodetraits}:
\begin{lstlisting}[language = sh,label={lst:readnodetraits-rho}]
readnodetraits --burnin 1000
               --var_within data/body_size/mammals.male.var_trait.tsv
               --output results_mammals_male.tsv
               run_mammals_male
\end{lstlisting}
The file \texttt{data\_empirical/chain\_name.ratio.tsv} then contains the posterior mean of the ratio of between-species variation over within-species variation, the 95\% and 99\% credible interval, and the posterior probability that the ratio is greater than 1.

\subsection{Maximum likelihood estimation}\label{subsec:maximum-likelihood-estimation}

To obtain the ratio (without the posterior credible interval and probability) using maximum likelihood computation, the following python script can be used:
\begin{lstlisting}[language = sh, label={lst:neutrality_index}]
python3 utils/neutrality_index.py --tree data/body_size/mammals.male.tree
                                  --traitsfile data/body_size/mammals.male.tsv
                                  --var_within data/body_size/mammals.male.var_trait.tsv
                                  --output results_ML_mammals_male.tsv
\end{lstlisting}
\end{document}
%TC:endignore