### 2024-03-05 manuscript/figureS2.tex
Added figures for revision

\documentclass[10pt]{beamer}
\geometry{margin=0pt, nohead, nofoot, paperwidth=1080pt,paperheight=375pt}
\usepackage{pdfpages}
\usepackage{graphicx}
\pdfinclusioncopyfonts=1
\usepackage{pgfplots,pgf}
\usepgfplotslibrary{fillbetween}
\usebackgroundtemplate{\tikz\node[opacity=0]{};}
\setbeamertemplate{footline}[frame number]{}
\setbeamertemplate{navigation symbols}{}
\setbeamertemplate{footline}{}
\usefonttheme{serif}

\begin{document}

    \centering
    \begin{minipage}{0.32\textwidth}
        \begin{center}
            \LARGE $\mu =\mu_{\text{ref}} / 2 $
        \end{center}
        \includegraphics[width=\textwidth, page=1] {artworks/distance.neutral.low_u.pdf}
    \end{minipage}
    \begin{minipage}{0.32\textwidth}
        \begin{center}
            \LARGE $\mu =\mu_{\text{ref}} \cdot 4 $
        \end{center}
        \includegraphics[width=\textwidth, page=1] {artworks/distance.neutral.high_u.pdf}
    \end{minipage}
    \begin{minipage}{0.32\textwidth}
        \begin{center}
            \LARGE $\mu =\mu_{\text{ref}} \cdot 20 $
        \end{center}
        \includegraphics[width=\textwidth, page=1] {artworks/distance.neutral.very_high_u.pdf}
    \end{minipage}
    \begin{minipage}{0.32\textwidth}
        \includegraphics[width=\textwidth, page=1] {artworks/pvalues.low_u.pdf}
    \end{minipage}
    \begin{minipage}{0.32\textwidth}
        \includegraphics[width=\textwidth, page=1] {artworks/pvalues.high_u.pdf}
    \end{minipage}
    \begin{minipage}{0.32\textwidth}
        \includegraphics[width=\textwidth, page=1] {artworks/pvalues.very_high_u.pdf}
    \end{minipage}

\end{document}